---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# unpivotr

[unpivotr](https://github.com/nacnudus/unpivotr) provides ools for converting
data from complex or irregular layouts to a columnar structure.  For example,
tables with multi-level column or row headers, or spreadsheets.  Header and data
cells are selected by their contents, position and formatting, and are
associated with one other by their relative positions.

Use [unpivotr](https://github.com/nacnudus/unpivotr) on Excel (.xlsx) files via
the [tidyxl](https://github.com/nacnudus/tidyxl) package.

## Installation

```{r, echo = TRUE, eval = FALSE}
devtools::install_github("nacnudus/tidyxl")
```

## Example

```{r, echo = TRUE}
library(unpivotr)
library(tidyxl)
library(readxl)
library(dplyr)
```

The package includes a spreadsheet, 'purpose.xlsx', which has tables that have
multi-row column headers and multi-column row-headers.  A popular package for
importing spreadsheets does the following:

```{r, echo = TRUE}
read_excel("./inst/extdata/purpose.xlsx", "NNW WNW")
```

The [tidyxl](https://github.com/nacnudus/tidyxl) package imports the same table
into a format suitable for [unpivotr](https://github.com/nacnudus/unpivotr):

```{r, echo = TRUE}
(pivoted <- contents("./inst/extdata/purpose.xlsx", "NNW WNW")[[1]])
```

We can identify each row of column headers, each column of row headers, and the
data cells, using any method.  Here, we use some
[dplyr](https://github.com/hadley/dplyr) and some base R functions.

```{r, echo = TRUE}
col_headers <- 
  pivoted %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers 

row_headers <- 
  pivoted %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

datacells <- 
  pivoted %>%
  filter(row >= 4, col >= 4) %>%
  select(row, col, value = as.integer(content))
datacells
```

Using [unpivotr](https://github.com/nacnudus/unpivotr) functions, we associate
the data cells with the headers, by proximity in given compass directions.

```{r, echo = TRUE}
unpivoted <- 
  datacells %>%
  NNW(col_headers[[1]], "sex") %>%
  N(col_headers[[2]], "purpose") %>%
  WNW(row_headers[[1]], "education") %>% 
  W(row_headers[[2]], "age")
unpivoted
```

We can re-pivot the final data in R using `ftable()` to check that it has been
imported correctly.

```{r, echo = TRUE}
ftable(unpivoted, 
       row.vars = c("education", "age"),
       col.vars = c("sex", "purpose"))
read_excel("./inst/extdata/purpose.xlsx", "NNW WNW")
```

The functions `extend()` and `offset` can be used when selecting cells relative
to an intial anchor cell (or group of cells, called a 'bag').

## Philosophy

Information in in many spreadsheets cannot be easily imported into R.  Why?

Most R packages that import spreadsheets have difficulty unless the layout of
the spreadsheet conforms to a strict definition of a 'table', e.g.:

* observations in rows
* variables in columns
* a single header row
* all information represented by characters, whether textual, logical, or
  numeric

These rules are designed to eliminate ambiguity in the interpretation of the
information.  But most spreadsheeting software relaxes these rules in a trade of
ambiguity for expression via other media:

* proximity (other than headers, i.e. other than being the first value at the
  top of a column)
* formatting (colours and borders)

Humans can usually resolve the ambiguities with contextual knowledge, but
computers are limited by their ignorance.  Programmers are hampered by:

* their language's expressiveness
* loss of information in transfer from spreadsheet to programming library

Information is lost when software discards it in order to force the data into
tabular form.  Sometimes date formatting is retained, but mostly formatting 
and comments are lost, and position has to be inferred.

## Similar projects

[unpivotr](https://github.com/nacnudus/unpivotr) is inspired by the United
Kingdom Office of National Statistics
[ONSdatabaker](https://github.com/ONS-OpenData/ONSdatabaker), which is
implemented in Python.  [unpivotr](https://github.com/nacnudus/unpivotr) differs
already by using compass directions, and further developments are planned.

The [rsheets](https://github.com/rsheets) project of several R packages is in
the early stages of importing spreadsheet information from Excel and Google
Sheets into R, manipulating it, and potentially parsing and processing formulas
and writing out to spreadsheet files.  In particular,
[jailbreaker](https://github.com/rsheets/jailbreakr) attempts to extract
non-tabular data from spreadsheets into tabular structures automatically via
some clever algorithms.

[unpivotr](https://github.com/nacnudus/unpivotr) differs from
[jailbreaker](https://github.com/rsheets/jailbreakr) in that it provides tools
for unpivoting complex and non-tabular data layouts using I not AI
(intelligence, not artificial intelligence).

## Roadmap

- [ ] Write tests
- [ ] Combine long header titles that are spread across cells.
