---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

# unpivotr

[![Travis-CI Build Status](https://travis-ci.org/nacnudus/unpivotr.svg?branch=master)](https://travis-ci.org/nacnudus/unpivotr)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/nacnudus/unpivotr?branch=master&svg=true)](https://ci.appveyor.com/project/nacnudus/unpivotr)
[![Cran Status](http://www.r-pkg.org/badges/version/unpivotr)](https://CRAN.R-project.org/package=unpivotr)
[![Cran Downloads](https://cranlogs.r-pkg.org/badges/unpivotr)](https://www.r-pkg.org/pkg/unpivotr)
[![codecov](https://codecov.io/github/nacnudus/unpivotr/coverage.svg?branch=master)](https://codecov.io/gh/nacnudus/unpivotr)

[unpivotr](https://github.com/nacnudus/unpivotr) deals with non-tabular data,
especially from spreadsheets.  Use unpivotr when your source data has any of
these 'features':

* Multi-headered hydra
* Meaningful formatting
* Headers anywhere but at the top of each column
* Non-text headers e.g. dates
* Other stuff around the table
* Several similar tables in one sheet
* Sentinel values
* Superscript symbols
* Meaningful comments
* Nested HTML tables

If that list makes your blood boil, you'll enjoy the function names.

* `behead()` deals with multi-headered hydra tables one layer of headers at a
  time, working from the edge of the table inwards.  It's a bit like using
  `header = TRUE` in `read.csv()`, but because it's a function, you can apply it
  to as many layers of headers as you need.  You end up with all the headers in
  columns.
* `spatter()` is like `tidyr::spread()` but preserves mixed data types.  You get
  into a mixed-data-type situation by delaying type coercion until *after* the
  table is tidy (rather than before, like `read.csv()` et al).  And yes, it
  usually follows `behead()`.

More positive, corrective functions:

* `justify()` aligns column headers before `behead()`ing, and has deliberate
  moral overtones.
* `enhead()` attaches a header to the body of the data, *a la* Frankenstein.
  The effect is the same as `behead()`, but is more powerful because you can
  choose exactly which header cells you want, paying attention to formatting
  (which `behead()` doesn't understand).
* `isolate_sentinels()` separates meaningful symbols like `"N/A"` or
  `"confidential"` from the rest of the data, giving them some time alone think
  about what they've done.
* `partition()` takes a sheet with several tables on it, and slashes into pieces
  that each contain one table.  You can then unpivot each table in turn with
  `purrr::map()` or similar.

## Getting started

[Spreadsheet Munging
Strategies](https://nacnudus.github.io/spreadsheet-munging-strategies) is a
free, online cookbook using [tidyxl](https://github.com/nacnudus/tidyxl) and
[unpivotr](https://github.com/nacnudus/unpivotr) (in progress).

Otherwise the basic idea is:

1. Read the data with a specialist tool.
   * For spreadsheets, use [tidyxl](https://nacnudus.github.io/tidyxl).
   * For plain text files, you might soon be able to use
     [readr](https://readr.tidyverse.org), but for now you'll have to install a
     pull-request on that package with
     `devtools::install_github("tidyverse/readr#760")`.
   * For tables in html pages, use `unpivotr::tidy_html()`
   * For data frames, use `unpivotr::tidy_table()` -- this should be a last
     resort, because by the time thee data is in a conventional data frame, it
     is often too late -- formatting has been lost, and most data types have
     been coerced to strings.
1. Either `behead()` straight away, else `dplyr::filter()` separately for the
   header cells and the data cells, and then recombine with `enhead()`.
1. `spatter()` so that each column has one data type.

```{r}
library(unpivotr)
library(tidyverse)
x <- purpose$`NNW WNW`
x # A pivot table in a conventional data frame.  Four levels of headers, in two
  # rows and two columns.

y <- tidy_table(x) # 'Tokenize' or 'melt' the data frame into one row per cell
y

rectify(y) # useful for reviewing the melted form as though in a spreadsheet

y %>%
  behead("NNW", "sex") %>%               # Strip headers
  behead("N", "life-satisfication") %>%  # one
  behead("WNW", "qualification") %>%     # by
  behead("W", "age-band") %>%            # one.
  select(-row, -col, -data_type, count = chr) %>% # cleanup
  mutate(count = as.integer(count))
```

Note the compass directions in the code above, which hint to `behead()` where to
find the header cell for each data cell.

* `"NNW"` means the header (`Female`, `Male`) is positioned up and to the
  left of the columns of data cells it describes.
* `"N"` means the header (`0 - 6`, `7 - 10`) is positioned directly above the
  columns of data cells it describes.
* `"WNW"` means the header (`Bachelor's degree`, `Certificate`, etc.) is
  positioned to the left and upwards of the rows of data cells it describes.
* `"W"` means the header (`15 - 24`, `25 - 44`, etc.) is positioned directly to
  the left of the rows of data cells it describes.


## Installation

```{r, echo = TRUE, eval = FALSE}
# install.packages("devtools") # If you don't already have devtools
devtools::install_github("nacnudus/unpivotr", build_vignettes = TRUE)
```

Don't use the CRAN version -- it has none of the fun in this README, and the
development version breaks everything.  Please let me know if this is a problem
for you --- as far as I know, nobody really used the old version, but it's
hard to tell.

## Similar projects

[unpivotr](https://github.com/nacnudus/unpivotr) is inspired by the United
Kingdom Office of National Statistics
[ONSdatabaker](https://github.com/ONS-OpenData/ONSdatabaker), which is
implemented in Python.  [unpivotr](https://github.com/nacnudus/unpivotr).

[jailbreaker](https://github.com/rsheets/jailbreakr) attempts to extract
non-tabular data from spreadsheets into tabular structures automatically via
some clever algorithms.  [unpivotr](https://github.com/nacnudus/unpivotr)
differs by being less magic, and equipping you to express what you want to do.
