---
title: "Compass Directions"
author: "Duncan Garmonsway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compase Directions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Compass directions

This vignette for the [unpivotr](https://github.com/nacnudus/unpivotr) package
demonstrates unpivoting pivot tables of various kinds imported from a
spreadsheet via the [tidyxl](https://github.com/nacnudus/tidyxl) package.

The spreadsheet is in the package directory at `../inst/extdata/purpose.xlsx`.

```{r, echo = TRUE}
library(dplyr)
library(tidyxl)
```

## NNW WNW example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "NNW WNW")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  NNW(col_headers[[1]], "sex") %>%
  N(col_headers[[2]], "purpose") %>%
  WNW(row_headers[[1]], "education") %>% 
  W(row_headers[[2]], "age")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

NNW_WNW <- datacells
```

## NNE WSW example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "NNE WSW")[[1]]
```

### Headers (same as NNW WNW)

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data (same as NNW WNW)

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
```

### Join headers to data (different from NNW WNW)

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  NNE(col_headers[[1]], "sex") %>% # Different from NNW WNW
  N(col_headers[[2]], "purpose") %>% # Same as NNW WNW
  WSW(row_headers[[1]], "education") %>% # Different from NNW WNW
  W(row_headers[[2]], "age") # Same as NNW WNW
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

NNE_WSW <- datacells
```

## SSE ESE example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "SSE ESE")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 8, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 24, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 23, col <= 7) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  SSE(col_headers[[2]], "sex") %>%
  S(col_headers[[1]], "purpose") %>%
  ESE(row_headers[[2]], "education") %>%
  E(row_headers[[1]], "age")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

SSE_ESE <- datacells
```

## SSW ENE example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "SSW ENE")[[1]]
```

### Headers (same as SSE ESE)

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 8, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 24, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data (same as SSE ESE)

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 23, col <= 7) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  SSW(col_headers[[2]], "sex") %>% # Different from SSE ESE
  S(col_headers[[1]], "purpose") %>% # Same as SSE ESE
  ENE(row_headers[[2]], "education") %>% # Different from SSE ESE
  E(row_headers[[1]], "age") # Same as SSE ESE
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

SSW_ENE <- datacells
```

## Check that all sheets are parsed correctly

```{r, echo = TRUE}
identical(NNW_WNW, NNE_WSW)
identical(NNW_WNW, SSE_ESE)
identical(NNW_WNW, SSW_ENE)
```

## ABOVE LEFT example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "ABOVE LEFT")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  ABOVE(col_headers[[1]], "purpose") %>% # Different from SSE ESE
  N(col_headers[[2]], "age") %>% # Same as SSE ESE
  LEFT(row_headers[[1]], "sex") %>% # Different from SSE ESE
  W(row_headers[[2]], "education") # Same as SSE ESE
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

ABOVE_LEFT <- datacells
```

## BELOW RIGHT example

```{r, echo = TRUE}
cells <- contents("../inst/extdata/purpose.xlsx", "BELOW RIGHT")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 10, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 14, !is.na(content)) %>%
  select(row, col, header = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 13, col <= 9) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  BELOW(col_headers[[2]], "purpose") %>%
  S(col_headers[[1]], "age") %>%
  RIGHT(row_headers[[2]], "sex") %>%
  E(row_headers[[1]], "education")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

BELOW_RIGHT <- datacells
```

## Check that all sheets are parsed correctly

```{r, echo = TRUE}
identical(ABOVE_LEFT, BELOW_RIGHT)
```

