---
title: "Compass Directions"
author: "Duncan Garmonsway"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Compase Directions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Compass directions

This vignette for the [unpivotr](https://github.com/nacnudus/unpivotr) package
demonstrates unpivoting pivot tables of various kinds imported from a
spreadsheet via the [tidyxl](https://github.com/nacnudus/tidyxl) package.  It is
best read with the spreadsheet open in a spreadsheet program, e.g. Excel,
LibreOffice Calc or Gnumeric.

The spreadsheet is in the package directory at `../inst/extdata/purpose.xlsx`.

```{r, echo = TRUE}
library(dplyr)
library(tidyxl)
library(unpivotr)

# Load spreadsheet of examples
spreadsheet <- system.file("extdata/purpose.xlsx", package = "unpivotr")
```

## NNW WNW example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "NNW WNW")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>% # Select all rows of headers at once
  select(row, col, value = character) %>%
  split(.$col) # Return each row of headers in its own element of a list
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  NNW(col_headers[[1]], "sex") %>%
  N(col_headers[[2]], "purpose") %>%
  WNW(row_headers[[1]], "education") %>% 
  W(row_headers[[2]], "age")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

NNW_WNW <- datacells
```

## NNE WSW example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "NNE WSW")[[1]]
```

### Headers (same as NNW WNW)

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data (same as NNW WNW)

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
```

### Join headers to data (different from NNW WNW)

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  NNE(col_headers[[1]], "sex") %>% # Different from NNW WNW
  N(col_headers[[2]], "purpose") %>% # Same as NNW WNW
  WSW(row_headers[[1]], "education") %>% # Different from NNW WNW
  W(row_headers[[2]], "age") # Same as NNW WNW
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

NNE_WSW <- datacells
```

## SSE ESE example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "SSE ESE")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 8, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 24, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 23, col <= 7) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  SSE(col_headers[[2]], "sex") %>%
  S(col_headers[[1]], "purpose") %>%
  ESE(row_headers[[2]], "education") %>%
  E(row_headers[[1]], "age")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

SSE_ESE <- datacells
```

## SSW ENE example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "SSW ENE")[[1]]
```

### Headers (same as SSE ESE)

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 8, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 24, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data (same as SSE ESE)

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 23, col <= 7) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  SSW(col_headers[[2]], "sex") %>% # Different from SSE ESE
  S(col_headers[[1]], "purpose") %>% # Same as SSE ESE
  ENE(row_headers[[2]], "education") %>% # Different from SSE ESE
  E(row_headers[[1]], "age") # Same as SSE ESE
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

SSW_ENE <- datacells
```

## Check that all sheets are parsed correctly

```{r, echo = TRUE}
identical(NNW_WNW, NNE_WSW)
identical(NNW_WNW, SSE_ESE)
identical(NNW_WNW, SSW_ENE)
```

## ABOVE LEFT example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "ABOVE LEFT")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  ABOVE(col_headers[[1]], "purpose") %>% # Different from SSE ESE
  N(col_headers[[2]], "age") %>% # Same as SSE ESE
  LEFT(row_headers[[1]], "sex") %>% # Different from SSE ESE
  W(row_headers[[2]], "education") # Same as SSE ESE
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

ABOVE_LEFT <- datacells
```

## BELOW RIGHT example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "BELOW RIGHT")[[1]]
```

### Headers

```{r, echo = TRUE}
row_headers <- 
  cells %>%
  filter(col >= 10, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 14, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
datacells <- 
  cells %>%
  filter(row <= 13, col <= 9) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  BELOW(col_headers[[2]], "purpose") %>%
  S(col_headers[[1]], "age") %>%
  RIGHT(row_headers[[2]], "sex") %>%
  E(row_headers[[1]], "education")
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

BELOW_RIGHT <- datacells
```

## Check that all sheets are parsed correctly

```{r, echo = TRUE}
identical(ABOVE_LEFT, BELOW_RIGHT)
```

## ABOVE LEFT border example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "ABOVE LEFT border")[[1]]
```

### Headers

```{r, echo = TRUE}
# Same as ABOVE LEFT without borders
row_headers <- 
  cells %>%
  filter(col <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row <= 3, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
# Same as ABOVE LEFT without borders
datacells <- 
  cells %>%
  filter(row >= 4, col >= 4) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Borders

```{r, echo = TRUE}
# Different from ABOVE LEFT without borders
# Find cells with borders on the bottom, and ones with borders on the left
formatting <- formats(spreadsheet)
left_borders <- which(!is.na(formatting$local$border$left$style))
top_borders <- which(!is.na(formatting$local$border$top$style))

left_border_cells <- 
  cells %>%
  filter(row == 2, local_format_id %in% left_borders) %>%
  select(row, col, value = content) # value doesn't matter
top_border_cells <- 
  cells %>%
  filter(col == 2, local_format_id %in% top_borders) %>%
  select(row, col, value = content) # value doesn't matter
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  ABOVE(col_headers[[1]], "purpose", left_border_cells) %>% # Different from ABOVE LEFT
  N(col_headers[[2]], "age") %>% # Same as ABOVE LEFT
  LEFT(row_headers[[1]], "sex", top_border_cells) %>% # Different from ABOVE LEFT
  W(row_headers[[2]], "education") # Same as ABOVE LEFT
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

ABOVE_LEFT_borders <- datacells
ABOVE_LEFT
```

## BELOW RIGHT border example

```{r, echo = TRUE}
cells <- contents(spreadsheet, "BELOW RIGHT border")[[1]]
```

### Headers

```{r, echo = TRUE}
# Same as BELOW RIGHT without borders
row_headers <- 
  cells %>%
  filter(col >= 10, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$col)
row_headers

col_headers <- 
  cells %>%
  filter(row >= 14, !is.na(content)) %>%
  select(row, col, value = character) %>%
  split(.$row)
col_headers
```

### Data

```{r, echo = TRUE}
# Same as BELOW RIGHT without borders
datacells <- 
  cells %>%
  filter(row <= 13, col <= 9) %>%
  mutate(content = ifelse(is.na(character), content, NA)) %>%
  select(row, col, value = as.integer(content))
datacells
```

### Borders

```{r, echo = TRUE}
# Different from BELOW RIGHT without borders
# Find cells with borders on the bottom, and ones with borders on the left
formatting <- formats(spreadsheet)
left_borders <- which(!is.na(formatting$local$border$left$style))
top_borders <- which(!is.na(formatting$local$border$top$style))

left_border_cells <- 
  cells %>%
  filter(row == 14, local_format_id %in% left_borders) %>%
  select(row, col, value = content) # value doesn't matter
top_border_cells <- 
  cells %>%
  filter(col == 11, local_format_id %in% top_borders) %>%
  select(row, col, value = content) # value doesn't matter
```

### Join headers to data

```{r, echo = TRUE}
datacells <- 
  datacells %>%
  BELOW(col_headers[[2]], "purpose", left_border_cells) %>% # Different from BELOW RIGHT
  S(col_headers[[1]], "age") %>% # Same as BELOW RIGHT
  RIGHT(row_headers[[2]], "sex", top_border_cells) %>% # Different from BELOW RIGHT
  E(row_headers[[1]], "education") # Same as BELOW RIGHT
```

### Result

```{r, echo = TRUE}
datacells %>% as.data.frame

BELOW_RIGHT_borders <- datacells
```

## Check that all sheets are parsed correctly

```{r, echo = TRUE}
identical(ABOVE_LEFT, BELOW_RIGHT)
identical(ABOVE_LEFT, ABOVE_LEFT_borders)
identical(ABOVE_LEFT, BELOW_RIGHT_borders)
```

